---
title: "Activity 3 - Bayesian Network 2"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---
###  Francisco Caio Lima Paiva - 5992344
**A)** Load adult and adult.text datasets with NA removed
```{r echo=FALSE}
library(bnlearn)
library(caret)
df = na.omit(read.csv("~/Documents/Poli/Mestrado/PCS5024-statML/adultbase/adult.data", na.strings="?", strip.white = TRUE))
df.test = na.omit(read.csv("~/Documents/Poli/Mestrado/PCS5024-statML/adultbase/adult.test", na.strings="?", strip.white = TRUE))
cat("adult train set dimensions: ", dim(df))
cat("\nadult test set dimensions: ", dim(df.test))
```

**B)** Choosed 5 features considered most important and reduced number of possible values for each feature to a factor of two
```{r echo=FALSE}
keeps = c("gross.income", "education.num", "marital.status", "sex", "age")
df_redux = df[keeps]
df_redux.test = df.test[keeps]

df_redux$marital.status = ifelse(df_redux$marital.status == "Married-civ-spouse","married", "not-married")
df_redux$education.num = ifelse(df_redux$education.num < 11, "below-bachelor", "equal-above-bachelor")
df_redux$age = ifelse(df_redux$age < 35 | df_redux$age > 64, "young-old", "prime-age")

df_redux.test$marital.status = ifelse(df_redux.test$marital.status == "Married-civ-spouse","married", "not-married")
df_redux.test$education.num = ifelse(df_redux.test$education.num < 11, "below-bachelor", "equal-above-bachelor")
df_redux.test$age = ifelse(df_redux.test$age < 35 | df_redux.test$age > 64, "young-old", "prime-age")

df_redux[] = lapply(df_redux, factor)
df_redux.test[] = lapply(df_redux.test, factor)
```

**C)** The following approaches to performing predictions were performed:

**1.**  Manually ajusted bayesian network structure and parameters.

**1.1.**  Bayesian Network manually designed:
```{r echo = FALSE}
arc.set=matrix(c("age", "gross.income", "education.num","gross.income","marital.status","gross.income","sex", "marital.status"),
               ncol=2,byrow=TRUE,
               dimnames=list(NULL,c("from","to")))

graph = empty.graph(keeps)

arcs(graph)=arc.set
plot(graph)
```

**1.2.**  Manually adjusting the parameters for the conditional probability tables (cpts). Ex:
```{r echo=FALSE}


cpt.gross_income = c(0.6, 0.4, 0.3, 0.7, 0.95, 0.05, 0.75, 0.25, 0.8, 0.2, 0.45, 0.55, 0.95, 0.05, 0.9, 0.1)
dim(cpt.gross_income) = c(2,2,2,2)
dimnames(cpt.gross_income) = list("gross.income" = c("<=50K", ">50K"), "education.num" = c("below-bachelor", "equal-above-bachelor"), "marital.status" = c("married", "not-married"), "age" = c("prime-age", "young-old"))

cpt.education_num = matrix(c(0.7, 0.3), ncol = 2, dimnames = list(NULL, c("below-bachelor", "equal-above-bachelor")))

cpt.marital_status = c(0.2, 0.8, 0.6, 0.4)
dim(cpt.marital_status) = c(2,2)
dimnames(cpt.marital_status) = list( "marital.status" = c("married", "not-married"), "sex" = c("Female", "Male"))

cpt.sex = matrix(c(0.3, 0.7), ncol = 2, dimnames = list(NULL, c("Female", "Male")))

cpt.age = matrix(c(0.6, 0.4), ncol = 2, dimnames = list(NULL, c("prime-age", "young-old")))

fitted = custom.fit(graph, dist = list(gross.income = cpt.gross_income, education.num= cpt.education_num, marital.status = cpt.marital_status, sex = cpt.sex, age = cpt.age))

print(fitted$marital.status)
```

**1.3.**  Finally attained the score below:
```{r echo=FALSE}
pred = predict(fitted, node="gross.income", data=df_redux.test, method = "bayes-lw", prob = TRUE)
xtab = table(pred, df_redux.test[, "gross.income"])
confmatman = confusionMatrix(xtab, positive = '>50K')
print(confmatman$table)
print(confmatman$overall['Accuracy'])
```

**2.**  Manually ajusted bayesian network structure and automatically obtained parameters.

**2.1.**  Using previous bayesian network, learned the parameters and obtained the conditional probability tables (cpts). Ex:
```{r echo=FALSE}
fitted = bn.fit(graph, data = df_redux)
print(fitted$marital.status)
```

**2.2.**  Finally, attained the score below:
```{r echo=FALSE}
pred = predict(fitted, node="gross.income", data=df_redux.test, method = "bayes-lw", prob = TRUE)
xtab = table(pred, df_redux.test[, "gross.income"])
confmatpartman = confusionMatrix(xtab, positive = '>50K')
print(confmatpartman$table)
print(confmatpartman$overall['Accuracy'])
```

**3.**  Automatically ajusted bayesian network structure and parameters.

**3.1.**  Learned the following structure using max-min hill climbing algorithm
```{r echo = FALSE}
res = hc(df_redux)
plot(res)
```

**3.2.**  Score below:
```{r echo=FALSE}
fitted = bn.fit(res, data = df_redux)
pred = predict(fitted, node="gross.income", data=df_redux.test, method = "bayes-lw", prob = TRUE)
xtab = table(pred, df_redux.test[, "gross.income"])
confmatauto = confusionMatrix(xtab, positive = '>50K')
print(confmatauto$table)
print(confmatauto$overall['Accuracy'])
```

**4.**  Naive Bayes

**4.1.**  Structure:
```{r echo = FALSE}
bn = naive.bayes(df_redux, "gross.income")
plot(bn)
```

**4.2.**  Score below:
```{r echo = FALSE}
fitted = bn.fit(bn, data = df_redux)
pred = predict(fitted, df_redux.test)
xtab = table(pred, df_redux.test[, "gross.income"])
confmatnb = confusionMatrix(xtab, positive = '>50K')
print(confmatnb$table)
print(confmatnb$overall['Accuracy'])
```

**D)** Final results are that approach 2 had marginally better accuracy than 1 and 3. Naive Bayes had the worst accuracy.
```{r echo= FALSE}
cat("Accuracy")
cat("\n Approach 1 - structure and cpts manual:", confmatman$overall['Accuracy'])
cat("\n Approach 2 - structure manual and cpts auto:", confmatpartman$overall['Accuracy'])
cat("\n Approach 3 - structure and cpts auto:", confmatauto$overall['Accuracy'])
cat("\n Approach 4 - naive bayes:", confmatnb$overall['Accuracy'])
```